Synthesizing Symmetric Lenses

### 2018-09-19 papers/SymmetricLenses/response_popl19.md

                                -------------

     - 10. p18, l842

           Could you explain in some more detail what criteria does `EXPAND` use
           to decide to open some closed expressions but not others?

At a high level, we process both regular expression trees, looking for
situations where a regular expression is a subcomponent of one format but
not the other, and expand such REs (this is where hash-consing provides its
greatest benefits). Furthermore, we look for instances where a regular
expression may be present in one format, but is only contained within a
closed regular expression in the other, and we expand the closed RE
containing the missing RE in this case. We then make this apply in slightly
more places by also taking into account how deeply nested within stars
certain REs occur.

@@

Both edit and update lenses would require new forms of specifications, since
both permit updates that are not merely pushing data through, but instead
are applying specific elements of an edit monoid.  (Also we would remove
complements from edit lenses).

===========================================================================
Review #43B

     - The main weakness is in the evaluation part. There are many 
       unclear/controversial parts regarding the evaluation. For example, the 
       lack of description of the benchmark environment; the way they test 
       whether the synthesised lens is correct; the description of the tasks to
       be synthesised. See detailed comments below. (This defect does not
       prevent it from being a good paper.)

We responded to the key points in this comment (the benchmark suite and
judging correctness) "above the fold" at the top of this response.

### 2018-09-19 papers/SymmetricLenses/response_popl19.md

1.  We manually inspected the code. This was certainly a nontrivial task, as
    these generated lenses are quite large, but perhaps easier than it may
    initially sound. Some of the lenses' structure follows from the structure of
    the regular expressions that describe source and target format --- this is
    relatively easy to validate. However, it can be more difficult to check the
    interactions of the complex combinators: Are the swaps in the right place?
    Are the correct components disconnected? Are lenses merged in the correct
    way? These complex combinators often occupied a few dozen lines of code
    rather than a few hundred, making our task easier. (Certainly, checking the
    code is easier than creating it de novo.)
2.  We built a few unit tests to validate the more complex lenses. 
3.  We ran the lenses on a number of strings, taken from the unit tests and the
    examples, but with minor alterations. We did not add these tests as
    additional unit tests, but they served as additional validation for
    ourselves as we were constructing the benchmark suite.

We recognize we did not go into significant enough detail on this in the
submitted version, and will gladly add in such detail on validation in the final
version.

                                -------------

      - There were a number of related questions about our information theoretic
        measure: Is it useful? Why not use a syntactic metric? Why does a
        fixed distribution work?  Will knowing the data distribution help?

Having a guiding principle behind these heuristics is, in itself, useful. Ad-hoc
syntactic metrics that are present in other systems (and we claim: are typically
building towards our underlying principle). Finding a simple underlying
principle (be as bijective as possible, which we formalize via information
theory) is better than a metric that simply applies only to our domain. We claim
that information theory is a general-purpose way to group all these metrics
where people give higher cost to their "less bijective" combinators (avoiding
constants, for example).

A fixed distribution works because, the core thing we are aiming for is "more
bijective." A fixed distribution is merely a way to not unfairly weight one
conjunct or disjunct over another in the presence of not knowing other
information. Even with our inferred distribution, we are still aiming towards
"more bijective".

Knowing the distribution will help in certain scenarios, and hinder in others.
With a learned distribution, our information theoretic measure is closer to the
ground-truth, permitting our learned lens to, in fact, be "more bijective."
However, we recognize that, while we believe "more bijective" is a solid guiding
principle, there are certain situations where the most bijective lens is not the
desired one: this is why we must further use examples for specifications. We
could certainly imagine situations where a learned distribution would find the
wrong lens (without additional examples), and synthesizing with the fixed
distribution would find the correct one, and vice-versa. However, we claim that,
as we typically want "more bijective," and a learned distribution gets closer to
a ground-truth on that, a learned distribution likely would perform better.

@@

     - 3. p5 l236,

          Shouldn't the second argument to `disconnect` be `company`?

We should have the second type be "" (to be consistent with what we use in the
final lens)

BCP: ?? 
AFM: In hindsight, I think I will just delete this question - it's more
of a "finding-errata" question than an actual one

@@

                                -------------

     - 9. p16, l777,

          Could you give concrete examples of situations where all sequences
          are involved in sequence lenses, but there are examples that could
          make other lenses more useful?

Within a DNF lens itself, all sequences are *required* to be involved with
another sequence lens: otherwise when that side of the DNF SRE is provided -
what disjunct on the other side should that lens be mapped to? Of course, there
are other ways the disjunct can be mapped: but these must happen at a different
level of the synthesis algorithm. For example - the entirety of that DNF regular
expression (which itself could be deeply nested within other DNF regular
expression components) could be projected. Furthermore, a specific sequence
could be mapped to another, but all the components of that sequence lens
actually may be disconnects. However, all sequences *must* be involved in a
sequence lens in a well-typed DNF lens.

BCP: I'm not sure I understand the question! 
AFM: I also now realize I didn't. After rereading the section, I think this
response is more appropriate.

@@

     - 10. p18, l842

           Could you explain in some more detail what criteria does `EXPAND` use
           to decide to open some closed expressions but not others?

At a high level, we process both regular expression trees, looking for
situations where a regular expression is as a subcomponent of one format but
not the other, and expand such REs (this is where hash-consing provides its
greatest benefits). Furthermore, we look for instances where a regular
expression may be present in one format, but is only contained within a
closed regular expression in the other, and we expand the closed RE
containing the missing RE in this case. We then make this apply in slightly
more places by also taking into account how deeply nested within stars
certain REs occur.

BCP: What is "closed"?
AFM: The internals of the Regular Expression are not exposed to GreedySynth,
they are the analogue of User Defined regular expressions in the previous paper,
but with this closed/open approach, we aren't dependent on where users place
variables (also making our integration into Boomerang significantly less hacky).

@@

All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16
GB of 1600 MHz DDR3 running macOS High Sierra.

                                -------------

     - The first subsection of Related Work did not give any reference to
       the related work except for Foster et al. 2007, which makes the
       paper unfriendly and hard to verify. The appendix is also in a
       messy.

We apologize for the lack of citations and the messy appendix. For the first
subsection of Related Work, we got carried away in making formal claims, and did
not include informal comparisons to other lens languages. We will make sure to
make such comparisons and citations in the next version of the paper. We will
similarly improve the appendix.

@@

We will make this more clear in future versions of the paper. Expand is
needed because GreedySynth alone cannot traverse regular expression
equivalences. If given the two regular expressions "" | "a"+, and "a"\*,
GreedySynth alone would merge "" to "a"\* (with a disconect), and merge "a"+
(also with a disconect).  However, contuing search through equivalent REs
helps find the bijective lens.

                                -------------

    - The ToStochastic procedure felt quite ad hoc to me. The paper makes a 
      point that if the user knew the distribution (e.g., if they automatically
      infer the parameters from data), then they can supply it. But will that
      really help the search? Essentially, I wish there were examples (even 
      synthetic ones) where knowing the distribution of data can clearly 
      influence the synthesis procedure.

We respond to this primarily above.  For a specific example:

Consider finding a lens of type:
("a" |(.5) "b") |(.5) "c" <=> "x" |(.1) "y"

In this situation, the lowest cost lens would merge "a" and "b" into "x", and
keep "c" mapped to "y".

Consider finding a lens of type:
("a" |(.5) "b") |(.5) "c" <=> "x" |(.9) "y"

In this situation, the lowest cost lens would merge "a" and "b" into "y", and
keep "c" mapped to "x".

We are happy to provide an example such as this one in the paper.

@@

We agree that section 7 requires more, informal comparisons to other lenses, for
us to discuss comparisons between our framework and other types of lenses. In
general, for different types of lenses to work, we require additional work,
either in the synthesis algorithm (for different types of combinators like
matching lenses or delta lenses), or in both the synthesis algorithm and
specifications (like for edit lenses and update lenses).

                                -------------

    - My main concern is whether the use of stochastic REs is really useful (see
      above)

Stochastic REs, and more generally, a semantic cost metric is critical for
comparing between responses by GreedySynth. Without this semantic cost, the same
function would likely have different costs if found through different calls to
GreedySynth. Syntactic metrics, we feel, aim to approximate how good the
underlying result is: in this work we are able to find a semantic metric that we
can be found using syntactic information.

### 2018-09-20 papers/SymmetricLenses/response_popl19.md

     - How do we determine what a "correct" lens is? Can we make this process a 
       bit more automatic and objective?

Beyond the format specifications and examples, we do not have specifications
that define the conversion functions. Hence, we do not have an automated way to
measure the correctness of our lenses (beyond checking that each synthesized
lens does indeed transform the given examples correctly). In this way, our work
is similar to other published synthesis-from-examples research, such as
Gulwani's work on FlashFill and related projects.

To check the correctness of our tool on our benchmarks, we did what
programmers usually do:

1.  We manually inspected the code. This was certainly a nontrivial task, as
    these generated lenses are quite large, but perhaps easier than it may
    initially sound. Some of the lenses' structure follows from the structure of
    the regular expressions that describe source and target format --- this is
    relatively easy to validate. However, it can be more difficult to check the
    interactions of the complex combinators: Are the swaps in the right place?
    Are the correct components disconnected? Are lenses merged in the correct
    way? These complex combinators often occupied a few dozen lines of code
    rather than a few hundred, making our task easier. (Certainly, checking the
    code is easier than creating it de novo.)
2.  We built unit tests to validate the more complex lenses. 
3.  We ran the lenses on a number of strings, taken from the unit tests and the
    examples, but with minor alterations. We did not add these tests as
    additional unit tests, but they served as additional validation for
    ourselves as we were constructing the benchmark suite.

KSF: Does 3 add anything??  We could delete it and save some words.

@@

We recognize we did not go into significant enough detail on this in the
submitted version, and will gladly add in such detail on validation in the final
version.

                                -------------

      - There were a number of related questions about our information theoretic
        measure: Is it useful? Why not use a syntactic metric? Why does a
        fixed distribution work?  Will knowing the data distribution help?

Having a principle guiding our heuristics is, in itself, useful as it
helps explain the efficacy of the heuristic.  A general-purpose
principle that cuts across multiple domains, (be as bijective as
possible, formalized via standard information theory) provides more
general insights than heuristics driven by a single domain. We claim
that information theory is a general-purpose way to group the many
metrics in which people assign higher costs to "less bijective"
combinators (avoiding constants, for example).  We believe that the ad
hoc syntactic metrics that are present in other systems are typically
building towards something like the information theoretic measure we
propose.

### 2018-09-20 papers/SymmetricLenses/response_popl19.md

Beyond the format specifications and examples, we do not have specifications
that define the conversion functions. Hence, we do not have an automated way to
measure the correctness of our lenses (beyond checking that each synthesized
lens does indeed transform the given examples correctly). In this way, our work
is similar to other published synthesis-from-examples research, such as
Gulwani's work on FlashFill and related projects.

To check the correctness of our tool on our benchmarks, we did what
programmers usually do:

1.  We manually inspected the code. This was certainly a nontrivial task, as
    these generated lenses are quite large, but perhaps easier than it may
    initially sound. Some of the lenses' structure follows from the structure of
    the regular expressions that describe source and target format --- this is
    relatively easy to validate. However, it can be more difficult to check the
    interactions of the complex combinators: Are the swaps in the right place?
    Are the correct components disconnected? Are lenses merged in the correct
    way? These complex combinators often occupied a few dozen lines of code
    rather than a few hundred, making our task easier. (Certainly, checking the
    code is easier than creating it de novo.)
2.  We built unit tests to validate the more complex lenses. 
3.  We ran the lenses on a number of strings, taken from the unit tests and the
    examples, but with minor alterations. We did not add these tests as
    additional unit tests, but they served as additional validation for
    ourselves as we were constructing the benchmark suite.

KSF: Does 3 add anything?? We could delete it and save some words.
AFM: I just wanted to give them full information about how we made sure the
lenses were correct, and this was part of my process, because that's their
primary fear. This comes from R2, who is the main person I want to make happy.

@@

                                -------------

     - 3. p5 l236,

          Shouldn't the second argument to `disconnect` be `company`?

Correct. However, in signsight, we should change the example such that the
second type is "" (to be consistent with what we use in the full lens)

@@

These benchmarks are using the full "SS" mode. The only difference between these
benchmarks and the full mode is that termination parameters are ignored, and
are instead required to use the 0, 25, and -25 parameters.

                                -------------

      - Second, from what I remember from the literature on (ordinary) symmetric
        lenses, they do not lend themselves very well to analysis in terms of 
        spans of asymmetric lenses. With simple symmetric lenses, do you think
        the situation is better now that you have omitted the complement part?

We are not experts in this area, but from our brief research we think that it
may be slightly better. From "Spans of Delta Lenses," p1, equivalence relations
are needed in both the span-presentation and set-based presentation of symmetric
lenses to address "'hidden' details such as their *complements*." Simple
symmetric lenses intentionally have no such details. However, there seem to be
other complexities related to lens laws, requiring restrictions like very
well-behavedness, as (from "Spans of Lenses", p1) "in the category whose
morphisms are lenses, what one might expect to be the pullback of lenses need
not satisfy the universal property of a pullback." We do not believe our
restriction helps obviate these difficulties.

@@

Within our synthesis algorithm, you are, in general, correct. However,
Boomerang exists beyond Optician, and permits lenses like id("a"|"a"). In
our extension to Boomerang, we do not remove this freedom.

                                -------------

     - L557 and below: Why some of the entropy is a single value while some
       of them seems ranging over an interval? If the interval is used to
       handle `merge_left`, then the entropy of all combinators should be an
       interval. Currently only `id`, `disconnect`, and `merge_left` yield
       intervals.

This was us trying to make our equations more readable, but we did not
sufficiently explain what our syntax meant. The calls to H(...) return pairs,
and when multiplied by constants, or added together, they act as follows:

@@

We will make sure to explain this more clearly in a final version of the paper.

                                -------------

     - L718 and L719: Different kinds of data definitely have different
       shapes. I cannot understand why assigning a fixed probability will fit
       in all situations. Could you explain it a bit?

See the main response "above the fold" at the top of this response.

@@

We will update the lens section of the related work to be more thorough on the
related work that we do not make formal statements about, and include proper
citations.

                                -------------

     - L1111 and L1112: I cannot figure out where does `x` below the
       "horizontal line" come from

This is true for any arbitrary x (as that x gets "overwritten" by the user
provided x')

@@

We will make this more clear in future versions of the paper. Expand is
needed because GreedySynth alone cannot traverse regular expression
equivalences. If given the two regular expressions "" | "a"+, and "a"\*,
GreedySynth alone would merge "" to "a"\* (with a disconect), and merge "a"+
(also with a disconect).  However, contuing search through equivalent REs
helps find the bijective lens.

                                -------------

    - The ToStochastic procedure felt quite ad hoc to me. The paper makes a 
      point that if the user knew the distribution (e.g., if they automatically
      infer the parameters from data), then they can supply it. But will that
      really help the search? Essentially, I wish there were examples (even 
      synthetic ones) where knowing the distribution of data can clearly 
      influence the synthesis procedure.

We respond to this primarily in the main response "above the fold" at the top of
this response. For a specific example:

@@

We respond to these issues in the main reponse.

                                -------------

    - Section 6 or 7 should discuss examples that are outside the expressivity 
      of the lens language and/or take too much time. It was hard for me to 
      infer what the limitations of the approach are.

We agree that section 7 requires more, informal comparisons to other lenses, for
us to discuss comparisons between our framework and other types of lenses. In
general, for different types of lenses to work, we require additional work,
either in the synthesis algorithm (for synthesizing different types of
combinators like matching lenses or delta lenses), or in both the synthesis
algorithm and specifications (like for edit lenses and update lenses).

### 2018-09-20 papers/SymmetricLenses/response_popl19.md

     - How do we determine what a "correct" lens is? Can we make this process a 
       bit more automatic and objective?

Beyond the format specifications and examples, we do not have specifications
that define the conversion functions. Hence, we do not have an automated way to
measure the correctness of our lenses (beyond checking that each synthesized
lens does indeed transform the given examples correctly). In this way, our work
is similar to other published synthesis-from-examples research, such as
Gulwani's work on FlashFill and related projects.

To check the correctness of our tool on our benchmarks, we did what
programmers usually do:

1.  We manually inspected the code. This was certainly a nontrivial task, as
    these generated lenses are quite large, but perhaps easier than it may
    initially sound. Some of the lenses' structure follows from the structure of
    the regular expressions that describe source and target format --- this is
    relatively easy to validate. However, it can be more difficult to check the
    interactions of the complex combinators, like swaps, disconnects, and
    merges. These complex combinators often occupied a few dozen lines of code
    rather than a few hundred, making our task easier. (Certainly, checking the
    code is easier than creating it de novo.)
2.  We built unit tests to validate the more complex lenses. 
3.  We ran the lenses on a number of strings, taken from the unit tests and the
    examples, but with minor alterations. We did not add these tests as
    additional unit tests, but they served as additional validation for
    ourselves as we were constructing the benchmark suite.

@@

                                -------------

      - There were a number of related questions about our information theoretic
        measure: Is it useful? Why not use a syntactic metric? Why does a
        fixed distribution work?  Will knowing the data distribution help?

Having a principle guiding our heuristics is, in itself, useful as it
helps explain the efficacy of the heuristic.  A general-purpose
principle that cuts across multiple domains, (be as bijective as
possible, formalized via standard information theory) provides more
general insights than heuristics driven by a single domain. We claim
that information theory is a general-purpose way to group the many
metrics in which people assign higher costs to "less bijective"
combinators (avoiding constants, for example).  We believe that the ad
hoc syntactic metrics that are present in other systems are typically
building towards something like the information theoretic measure we
propose.

A fixed distribution works because the core thing we are aiming for is "more
bijective." A fixed distribution is merely a way to not unfairly weight one
conjunct or disjunct over another in the absence of additional information.
However, even with our fixed distribution, we are still aiming towards "more
bijective".

Knowing the distribution will help in certain scenarios, and hinder in others.
With a learned distribution, our information theoretic measure is closer to the
ground-truth, permitting our learned lens to be "more bijective." However, while
we believe "more bijective" is a solid guiding principle, there are certain
situations where the most bijective lens is not the desired one: this is why we
must further use examples for specifications. We could certainly imagine
situations where a learned distribution would not find the desired lens (without
additional examples), and synthesizing with the fixed distribution would find
the desired one, and vice-versa. However, we claim that, as we typically want
"more bijective," and a learned distribution is closer to a ground-truth, a
learned distribution likely would perform better.

### 2018-09-20 papers/SymmetricLenses/response_popl19.md

1.  We manually inspected the code. This was certainly a nontrivial task, as
    these generated lenses are quite large, but perhaps easier than it may
    initially sound. Some of the lenses' structure follows from the structure of
    the regular expressions that describe source and target format --- this is
    relatively easy to validate. However, it can be more difficult to check the
    interactions of the complex combinators, like swaps, disconnects, and
    merges. These complex combinators often occupied a few dozen lines of code
    rather than a few hundred, making our task easier. (Certainly, checking the
    code is easier than creating it de novo.)
2.  We ran a set of tests to validate the more complex lenses.

We recognize we did not go into significant enough detail on this in the
submitted version, and will gladly add in such detail on validation in the final
version.

                                -------------

      - There were a number of related questions about our information theoretic
        measure: Is it useful? Why not use a (complex) syntactic metric? Why does a
        fixed distribution work?  Will knowing the data distribution help?

It is difficult (impossible) to prove that our information-theoretic
measure is better than all possible "syntactic" metrics, but our
experiments show that it does do well.  In particular, it outperforms
simple syntactic heuristics.  More important though, is the fact that
our heuristic is based on a relatively simple, clear idea that is
relatively easy to communicate and implement.  This idea can
potentially be reused in a variety of related settings (and then
perhaps augmented with ad hoc, domain-specific heuristics in those
settings).  In contrast, complex ad hoc syntactic metrics provide
little value beyond the single artifact in which they are used.

In addition, we believe that our information-theoretic ideas have
value because they help us understand some of the ad hoc metrics used in
the past: those in which people assign higher costs to "less
bijective" combinators in simpler ways (avoiding constants, for
example).  We believe that the ad hoc syntactic metrics that are
present in other systems are often building towards something like
the information theoretic measure we propose.

A fixed distribution works because the core goal is to generate "more
bijective" transformations.  A fixed distribution avoids unfairly
weighting one conjunct or disjunct over another in the absence of
additional information.  However, even with our fixed distribution, we
are still aiming towards "more bijective".

Knowing the distribution will help in certain scenarios, as it could
help to disambiguate between several possible "equally bijective"
lenses.  More broadly, the weightings can provide evidence that
data from one (heavily-weighted) component on the left flows to
another (heavily-weighted) component on the right.  For example:

ANDERS:

ANDERS POSSIBLY CUT THIS PARAGRAPH.  (I don't think it is detailed
enough/concrete enough to explain exactly how a weighted distribution
helps/hurts.):

### 2018-09-20 papers/SymmetricLenses/response_popl19.md

      - There were a number of related questions about our information theoretic
        measure: Is it useful? Why not use a (complex) syntactic metric? Why does a
        fixed distribution work?  Will knowing the data distribution help?

It is difficult (impossible) to prove that our information-theoretic
measure is better than all possible "syntactic" metrics, but our
experiments show that it does do well.  In particular, it outperforms
simple syntactic heuristics.  More important though, is the fact that
our heuristic is based on a relatively simple, clear idea that is
relatively easy to communicate and implement.  This idea can
potentially be reused in a variety of related settings (and then
perhaps augmented with ad hoc, domain-specific heuristics in those
settings).  In contrast, complex ad hoc syntactic metrics provide
little value beyond the single artifact in which they are used.

In addition, we believe that our information-theoretic ideas have
value because they help us understand some of the ad hoc metrics used in
the past: those in which people assign higher costs to "less
bijective" combinators in simpler ways (avoiding constants, for
example).  We believe that the ad hoc syntactic metrics that are
present in other systems are often building towards something like
the information theoretic measure we propose.

A fixed distribution works because the core goal is to generate "more
bijective" transformations --- the intuition is that information
should be preserved, when possible, as opposed to thrown away, when
data is translated between formats.  A fixed, uniform distribution avoids
unfairly weighting one conjunct or disjunct over another and unfairly
throwing certain information away.  In the absence of additional
information, this kind of fairness seems to be the best we can do.

On the other hand, knowing the distribution can help the system make a
choice about what to throw away --- given a choice, it prefers to
throw away less data.  Assume (x?.9) means "x appears
with probability .9 and is "" otherwise.  Consider a conversion between
these formats:

(x?.999).(y?.001) <==> z

Here, x appears more frequently than y on the left.  Hence, the system will prefer
a lens that projects y and maps x to z over a lens that projects x and
maps y to z as the latter tends to throw away much more information
and is "less bijective" in an information theoretic sense.

### 2018-09-20 papers/SymmetricLenses/response_popl19.md

BCP: The first section is too long -- 500 words max.

Thank you for your comments!

We begin with brief responses to the most significant concerns, followed by
an (optional) appendix addressing individual questions point by point.

@@

                                -------------

     - How do we determine what a "correct" lens is? Can we make this process a 
       bit more automatic and objective?

We agree we did not go into enough detail on this, and we can add more in a
final version.

We do not have an automated way to measure the correctness of our lenses
(beyond checking that each synthesized lens does indeed transform the given
examples correctly). In this respect, our work is similar to other published
synthesis-from-examples research, such as Gulwani's work on FlashFill and
related projects.

To assess the behavior of our tool on our benchmarks, we did what
programmers usually do:

1.  We manually inspected the code.

    This was a nontrivial task, as generated lenses can be quite large, but
    it was easier than it might initially sound. Some of the lenses'
    structure should follow from the structure of the regular expressions
    describing source and target formats; this is relatively easy to
    validate. The more difficult part was checking the interactions of
    complex combinators like swaps, disconnects, and merges. Still these
    bits typically spanned only a few dozen lines of code, making our task
    easier. Certainly, checking the code is easier than creating it de novo.

2.  We ran a set of tests to validate the more complex lenses.

BCP: #2 needs a few more words

@@

                                -------------

      - A number of related questions about our information-
        theoretic measure: Is it useful? Why not use a 
        (perhaps complex) syntactic metric? 
        Why does a fixed distribution work?  
        Will knowing the data distribution help?

BCP: Difficult, or impossible???

It is difficult (impossible) to prove that our information-theoretic measure
is better than all possible "syntactic" metrics, but our experiments show
that it performs well.  In particular, it outperforms simple syntactic
heuristics.  More important is the fact that our heuristic is based on a
simple, clear idea that is easy to communicate and implement.  This idea can
potentially be reused in a variety of related settings (and perhaps
_augmented_ with domain-specific heuristics).

BCP: I'd delete this: "By contrast, ad hoc syntactic metrics provide little
value beyond the single artifact in which they are used."

Moreover, our information-theoretic presentation has helped us understand
some of the ad hoc metrics used in the past -- [BCP: Whose?] those in which
people assign higher costs to "less bijective" combinators in simpler ways
(avoiding constants, for example).  These ad hoc syntactic metrics seem to
be approximating something like the information-theoretic measure we
work with.

A fixed distribution works because the core goal is to generate "more
bijective" transformations; the intuition is that information should be
preserved, when possible, when data is translated between formats.  A fixed,
uniform distribution avoids unfairly weighting one conjunct or disjunct over
another and unfairly throwing information away.  In the absence of
additional information, this kind of fairness seems to be the best we can
do.  [BCP: This is not very satisfying.  Seems like the question is exactly
"What kind of additional information might be useful and how could it be
communicated?"  Perhaps we could comment that this could be a direction for
future work.]

BCP: Oh, maybe the next paragraph is the discussion I was just asking for.
But I'm confused by the way it's introduced: Is this something we do now, or
something we could do?

On the other hand, knowing the distribution can help the system make a
choice about what to throw away: given a choice, it prefers to throw away
less data.  For example, let (x?.9) mean "x appears with probability .9 and
is "" otherwise and consider a conversion between these formats:

@@

    (x?.999).(y?.001) <==> z

Here, x appears more frequently than y on the left.  Hence, the system will
prefer a lens that projects y and maps x to z over a lens that projects x
and maps y to z, as the latter tends to throw away much more information and
is "less bijective" in an information-theoretic sense.

===========================================================================
===========================================================================
End of main response.

Detailed responses to individual questions in the reviews follow...

@@

                                -------------

     - 3. p5 l236,

          Shouldn't the second argument to `disconnect` be `company`?

Correct. However, in hindsight, we should change the example such that the
second type is "" (to be consistent with what we use in the full lens)

@@

                                -------------

     - 9. p16, l777,

          Could you give concrete examples of situations where all sequences
          are involved in sequence lenses, but there are examples that could
          make other lenses more useful?

Within a DNF lens itself, all sequences are *required* to be involved with
another sequence lens: otherwise, when that side of the DNF SRE is provided,
what disjunct on the other side should that lens be mapped to? Of course,
there are other ways the disjunct can be mapped, but these must happen at a
different level of the synthesis algorithm. For example, the entirety of
that DNF regular expression (which itself could be deeply nested within
other DNF regular expression components) could be projected. Furthermore, a
specific sequence could be mapped to another, but all the components of that
sequence lens actually may be disconnects. However, all sequences *must* be
involved in a sequence lens in a well-typed DNF lens.

@@

                                -------------

     - 10. p18, l842

           Could you explain in some more detail what criteria does `EXPAND` use
           to decide to open some closed expressions but not others?

At a high level, we process both regular expression trees, looking for
situations where a regular expression is a subcomponent of one format but
not the other, and expand such REs (this is where hash-consing provides its
greatest benefits). Also, we look for instances where a regular expression
may be present in one format, but is only contained within a "closed"
regular expression in the other, and we expand the closed RE containing the
missing RE in this case. We make this apply in a few more places by taking
into account how deeply nested within stars certain REs occur.

@@

BCP: What is "closed"?
AFM: The internals of the Regular Expression are not exposed to GreedySynth,
they are the analogue of User Defined regular expressions in the previous paper,
but with this closed/open approach, we aren't dependent on where users place
variables (also making our integration into Boomerang significantly less hacky).

BCP: OK.  Can a short version of that make it into a parenthesis after
"closed" above?

@@

                                -------------

     - L922: The authors may add a few words on how representative and
       complex the selected tasks is. For example, I got the data from the
       paper *Deep API Programmer: Learning to Program with APIs* that there
       is 238 real-world FlashFill benchmarks; then why the 8 data cleaning
       tasks here are more representative among the 238 tasks? The reviewers
       (and readers in the future) may want to know to which extent the job
       of writing lenses can be automatically done.

We chose our benchmarks from the FlashFill paper because the 238 tasks are,
unfortunately, not publicly available. In general, our benchmark suite goes
beyond the level of difficulty required for data cleaning. Many real-life
file format descriptors involve large numbers of disjunctions and nested
iterations, which FlashFill is either bad at (for disjunctions) or cannot
handle at all (nested iterations).

A more thorough comparison to FlashFill can be found in the previous
Synthesizing Bijective Lenses paper.

@@

    - The usability aspect is questionable: It seems that if the user has 
      defined complex regular expressions defining the pair of languages, then
      they can probably go ahead and just implement the lens without resorting 
      to examples.

As a demonstration about the difficulty of writing lenses, consider the
following bidirectional program

     let single_author_convert =
       ins " au - "
         . lens_swap
             (NAME . del ",")
             (lens_swap WSP NAME)*
     in ...

Even the relatively simple lens that permutes the 3 elements needed for
single_author_convert is fairly complex.  The complexity becomes even more
apparent when working with large lenses built from many sub-lenses.  Lenses
provide great power with their invertibility guarantees, but they come at
the cost of thinking about fiddly details when writing the terms.  Worrying
about these fiddly details while ALSO thinking about unambiguity
restrictions is a difficult task!

Engineering best practices suggest that it's a good idea to annotate lenses
with their types.  While Boomerang does not require these annotations (it
can infer the types from the terms), types serve as documentation for future
programmers to understand what formats the lens maps between. Furthermore,
the types provide resilience in the face of future lens modifications,
ensuring that the lens maps exactly between strings of the provided formats.

@@

                                -------------

    - The ToStochastic procedure felt quite ad hoc to me. The paper makes a 
      point that if the user knew the distribution (e.g., if they automatically
      infer the parameters from data), then they can supply it. But will that
      really help the search? Essentially, I wish there were examples (even 
      synthetic ones) where knowing the distribution of data can clearly 
      influence the synthesis procedure.

We respond to this primarily in the main response "above the fold" at the
top of this response. For a specific example:

Consider finding a lens of type:

    ("a" |(.5) "b") |(.5) "c" <=> "x" |(.1) "y"

@@

In this situation, the lowest cost lens would merge "a" and "b" into "x", and
keep "c" mapped to "y".

Consider finding a lens of type:

    ("a" |(.5) "b") |(.5) "c" <=> "x" |(.9) "y"

@@

We agree that section 7 requires more, informal comparisons to other lenses,
for us to discuss comparisons between our framework and other types of
lenses. In general, for different types of lenses to work, we require
additional work, either in the synthesis algorithm (for synthesizing
different types of combinators like matching lenses or delta lenses), or in
both the synthesis algorithm and specifications (like for edit lenses and
update lenses).

                                -------------

    - My main concern is whether the use of stochastic REs is really useful (see
      above)

Stochastic REs, and more generally, a semantic cost metric are critical for
comparing between responses by GreedySynth. Without this semantic cost, the
same function would likely have different costs if found through different
calls to GreedySynth. Syntactic metrics, we feel, aim to approximate how
good the underlying result is: in this work we are able to find a semantic
metric that we can be found using syntactic information.

@@

                                -------------

    - Are stochastic REs new? I didn't see any citations when they were 
      introduced.

Stochastic REs are not new. We cite them when we first mention them on page
2, but agree we should provide additional citations when they are formally
introduced. We do believe our syntactic technique for calculating their
entropy is novel.

### 2018-09-20 papers/SymmetricLenses/response_popl19.md

     - 10. p18, l842

           Could you explain in some more detail what criteria does `EXPAND` use
           to decide to open some closed expressions but not others?

At a high level, we process both regular expression trees, looking for
situations where a regular expression is a subcomponent of one format but
not the other, and expand such REs (this is where hash-consing provides its
greatest benefits). Furthermore, we look for instances where a regular
expression may be present in one format, but is only contained within a
closed regular expression in the other, and we expand the closed RE
containing the missing RE in this case. We then make this apply in slightly
more places by also taking into account how deeply nested within stars
certain REs occur.

Full details can be found in the prior work on Synthesizing Bijective Lenses
(page 16), though in that context, "user-defined data types" are used instead of
"closed regular expressions."

@@

Both edit and update lenses would require new forms of specifications, since
both permit updates that are not merely pushing data through, but instead
are applying specific elements of an edit monoid.  (Also we would remove
complements from edit lenses).

===========================================================================
Review #43B

     - The main weakness is in the evaluation part. There are many 
       unclear/controversial parts regarding the evaluation. For example, the 
       lack of description of the benchmark environment; the way they test 
       whether the synthesised lens is correct; the description of the tasks to
       be synthesised. See detailed comments below. (This defect does not
       prevent it from being a good paper.)

We responded to issues on judging correctness "above the fold" at the top of
this response.

@@

We will make this more clear in future versions of the paper. Expand is
needed because GreedySynth alone cannot traverse regular expression
equivalences. If given the two regular expressions "" | "a"+, and "a"\*,
GreedySynth alone would merge "" to "a"\* (with a disconect), and merge "a"+
(also with a disconect).  However, contuing search through equivalent REs
helps find the bijective lens.

                                -------------

    - The ToStochastic procedure felt quite ad hoc to me. The paper makes a 
      point that if the user knew the distribution (e.g., if they automatically
      infer the parameters from data), then they can supply it. But will that
      really help the search? Essentially, I wish there were examples (even 
      synthetic ones) where knowing the distribution of data can clearly 
      influence the synthesis procedure.

We respond to this primarily in the main response "above the fold" at the top of
this response.

We are happy to provide an example such as the one above in the paper.

### 2018-09-20 papers/SymmetricLenses/response_popl19.md

Both edit and update lenses would require new forms of specifications, since
both permit updates that are not merely pushing data through, but instead
are applying specific elements of an edit monoid.  (Also we would remove
complements from edit lenses).

===========================================================================
Review #43B

     - The main weakness is in the evaluation part. There are many 
       unclear/controversial parts regarding the evaluation. For example, the 
       lack of description of the benchmark environment; the way they test 
       whether the synthesised lens is correct; the description of the tasks to
       be synthesised. See detailed comments below. (This defect does not
       prevent it from being a good paper.)

We responded to issues on judging correctness "above the fold" at the top of
this response.

A more thorough comparison to FlashFill is present in the previous Synthesizing
Bijective Lenses paper.

### 2018-09-20 papers/SymmetricLenses/response_popl19.md

BCP: #2 needs a few more words

                                -------------

      - A number of related questions about our information-
        theoretic measure: Is it useful? Why not use a 
        (perhaps complex) syntactic metric? 
        Why does a fixed distribution work?  
        Will knowing the data distribution help?

It is impossible to prove that our information-theoretic measure
is better than all possible "syntactic" metrics, but our experiments show
that it performs well.  In particular, it outperforms simple syntactic
heuristics.  More important is the fact that our heuristic is based on a
simple, clear idea that is easy to communicate and implement.  This idea can
potentially be reused in a variety of related settings (and perhaps
_augmented_ with domain-specific heuristics).
